---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data

First unzip and read the CSV file into a variable.

The fields are:

+ **steps**: # of steps taken

+ **date**: in YYYY-MM-DD format

+ **interval**: 5 minute interval

```{r}
unzip("activity.zip")
data <- read.csv("activity.csv")
head(data)
summary(data)
```

## What is mean total number of steps taken per day?

First aggregate the data and plot a histogram of the total steps per day.  Then compute the mean and median of steps per day.  It would appear that people who track their steps walk a lot farther than the average person.

```{r}
aggdata <- aggregate(data$steps, by=list(data$date), FUN=sum, na.rm=T)
names(aggdata) <- c("date","steps")
hist(aggdata$steps, breaks=25, main="Histogram of Total Steps Per Day", xlab="Steps per Day",col=c("cadetblue1","aquamarine"))
steps.mean <- mean(aggdata$steps)
steps.median <- median(aggdata$steps)
print(paste( "Mean Steps: ", as.integer(steps.mean),"  Median Steps: ",steps.median))
```


## What is the average daily activity pattern?

```{r}
aggdata <- aggregate(data$steps, by=list(data$interval), FUN=mean, na.rm=T)
names(aggdata) <- c("interval","steps")

```


## Imputing missing values

```{r}
d2 <- na.omit(data)
```


## Are there differences in activity patterns between weekdays and weekends?

```{r}
```

